Embedded Scalars: {A}n Experimental Revisit

### 2014-05-15 text/paper/paper4.tex
update section 3

\usepackage[natbib=true,
            style=authoryear-comp,
            backend=bibtex,
            sortcites=false,
            maxnames=2,
            doi=false,
            url=false]{biblatex}
% \bibliography{../helpers/MyRefGlobal}
\bibliography{paper}

@@

Consequently, the \mymark{Gricean recipe}
\citep[see][]{Geurts2010:Quantity-Implic} for deriving a simple
scalar inference like that in (\ref{bsp:Plain-SI-Implicature}) from an
utterance of (\ref{bsp:Plain-SI-Target}) is as follows: if the issue
whether Hans solved only some or all of the problems is relevant, then
a cooperative and knowledgable speaker would utter
(\ref{bsp:Plain-SI-Alternative}) if in a position to do so; hence, one
of the most natural explanations of why such a speaker has not uttered
(\ref{bsp:Plain-SI-Alternative}), but only (\ref{bsp:Plain-SI-Target})
is that she is uncertain of whether (\ref{bsp:Plain-SI-Alternative})
is true; but on the assumption that she is knowledgeable (competent,
opinionated, informed \dots) it follows that
(\ref{bsp:Plain-SI-Implicature}) should in fact be true.\fn{We are
  glossing here somewhat swiftly over the more nuanced details of the
  derivation of implicatures targeting the speaker's epistemic state
  \citep[e.g.][]{Gazdar1979:Pragmatics:-Imp,Soames1982:How-Presupposit},
  as this is not crucially relevant for the issues we are interested
  in here.}

@@

Which of these readings, if any, does traditionalism predict to be
preferred? This depends on whether the auxiliary assumptions necessary
to derive global readings by the Gricean recipe are plausibly met in
the particular case of utterance of \as- and \es-sentences. These
auxiliary assumptions include relevance of the extra information
provided by the global reading, mutual awareness that the stronger
alternative has been a speaker option, the speaker's competence about
the issue, etc. Normally, traditionalist accounts would assume that
these extra assumptions are met. In that case, traditionalism would
predict that global readings should be preferred over literal
readings. On the other hand, it might also be hypothesized that, for
example, a competence assumption is harder to justify for \as- and
\es-sentences in general than for simpler sentences such as
(\ref{bsp:Plain-SI}): it might be less clear that the speaker knows
exactly how many students solved how many problems, than that the
speaker knows exactly how many problems Hans solved. In that case, or
if any other assumption of the Gricean recipe cannot be maintained,
traditionalism would predict that the literal reading would be
preferred over the global one. But that means that there are at least
two varieties of traditionalism that, depending on which additional
assumptions we would make, yield slightly different predictions:
\mymark{the strong variety of traditionalism} maintains that the
auxiliary assumptions of the Gricean recipe hold
usually/strongly/unless-completely-intenable and therefore predicts
that global readings are preferred over literal ones; \mymark{the weak
  variety of traditionalism} holds that the auxiliary assumptions are
more fragile and predicts that literal readings are preferred over the
global ones.\fn{We mention for clarity that although weak
  traditionalism does not predict the global reading to be preferred,
  it might still predict an \emph{epistemically weak implicature},
  similar to the global reading, that the speaker is uncertain whether
  the stronger alternative is true. Whether it does predict that
  depends on which auxiliary assumptions are assumed to be met and
  which are not.}

Does traditionalism also predict local readings to be available? The
answer is slightly different for \as- and the \es-sentences.\fn{Some
  traditionalists concede that (something like) local readings can
  occur for both kinds of sentences if scalar items are marked with
  special accentuation, albeit then as a signal of a different
  pragmatic process
  \citep[e.g.][]{Horn2006:The-Border-Wars,Geurts2009:Scalar-Implicat,Geurts2010:Quantity-Implic}. We
  will address this addition to traditionalism in more detail in
  Section~\ref{sec:inton-mark-hypoth}.} For \es-sentences,
traditionalism does not predict that a local reading is available, at
least not as a quantity implicature
\citep[see][]{GeurtsPouscoulous2009:Embedded-Implic,ChemlaSpector2010:Experimental-Ev}. This
is because traditionalism assumes that quantity implicatures are
pragmatic enrichments of the literal meaning of an utterance, obtained
by conjoining the literal meaning with a suitable set of negated
alternatives. But since the literal and the local reading of
\es-sentences are logically independent, there is no way that local
readings can be derived in a traditionalist manner as a quantity
implicature: if $X$ and $Y$ are logically independent propositions,
then there is no proposition $Z$ such that $X$ would be equivalent to
$Y \wedge \neg Z$.

@@

Another prominent theoretical position is
\mymark{grammaticalism}. Recently, a lot of evidence based on
intuitive judgements and more theoretical considerations has been
discussed in favor of this position
\citep[see][]{Chierchia2006:Broaden-Your-Vi,Fox2007:Free-Choice-and,Magri2011:Another-Argumen,Sauerland2012:The-Computation,ChierchiaFox2008:The-Grammatical,Chierchia2012:FC-Nominals-and}. % When
% it comes to the predictions of available readings of \as- and
% \es-sentences, grammaticalism appears like the conjunction of
% traditionalism and conventionalism. But conceptually speaking,
% grammaticalism goes down a road of its own.
Grammaticalism approaches quantity implicatures by postulating a
silent operator that can be variably applied during compositional
computation of a sentence's truth-value
\citep{Chierchia2006:Broaden-Your-Vi}, if necessary multiple times
\citep{Fox2007:Free-Choice-and}. This silent operator is variably
referred to as $\mathrm{O}(\cdot)$ or $\exh(\cdot)$, because it is
assumed to be similar in effect to the meaning of particle \emph{only}
or of the mechanism of \emph{exhaustive interpretation}
\citep{GroenendijkStokhofThesis1984,Stechowvon-StechowZimmermann1984:Term-Answers-an,Rooijvan-RooijSchulz2013:Exhaustive-Inte,vanRooijSchulz:ExhaustiveInterpretation,Fox2007:Free-Choice-and}. For
our purposes, it is enough to note that $\exh(\cdot)$ is a
poly-typed function that enriches an expression $X$, which crucially
need not be a full proposition, based on a set $\alt(X)$ of
(suitable, relevant) alternatives to $X$ that yields an enriched
meaning of the form:\fn{More sophisticated formulations of exhaustive
  interpretation have been proposed
  \citep[e.g.][]{Schulz2005:A-Pragmatic-Sol,vanRooijSchulz:ExhaustiveInterpretation,Spector2006:Scalar-Implicat,Fox2007:Free-Choice-and}
  but this simple formulation is sufficient for the purposes of this
  paper. Also, we gloss here over non-trivial details in the way this
  operation is to be specified exactly within a compositional
  semantics \citep[see][]{Chierchia2006:Broaden-Your-Vi}.}
\begin{exe}
  \ex \label{bsp:Exh-Def} $\exh(X,\alt(X)) = X \bigwedge_{A \in
      \alt(X)} \neg A$.
\end{exe}

@@

The grammatical approach, as described so far, is rather flexible in
that it makes many readings available. It is so flexible, in fact,
that it must be constrained in some way or other to shield the
approach against overgeneration. Grammaticalism is therefore aided by
a principle that specifies which readings are to be preferred if
several readings are generated by optional applications of
$\exh(\cdot)$ in the compositional computation of semantic values. The
principle that grammaticalists adhere to
\citep[see][]{FoxSpector:Economy-and-Emb,ChierchiaFox2008:The-Grammatical,Chierchia2012:FC-Nominals-and}
is the \mymark{strongest meaning hypothesis} of
\citet{DalrympleKanazawa1998:Reciprocal-Expr}. Roughly speaking, if a
given sentence has several conceivable readings, the strongest meaning
hypothesis selects for the strongest of
these. \citet{ChierchiaFox2008:The-Grammatical} discuss two concrete
variants of the strongest meaning hypothesis: fix a sentence with
propositional content $S$ with candidate readings $C(S)$, where $C(S)$
contains $S$ and all readings derivable from inserting $\exh(\cdot)$
at suitable scope sites;\fn{Strictly speaking, we should compare
  \emph{parses} of sentences with $\exh(\cdot)$ at various scope sites
  with respect to the readings that these parses give rise to
  \citep[see][]{ChierchiaFox2008:The-Grammatical}, but we may ignore
  this detail here for ease of exposition.} then define \emph{ceteris
  paribus} preferences among members of $X,Y \in C(S)$ as either
\begin{exe}
\ex \label{bsp:SMH-Total} 
  $X$ is preferred to $Y$ iff $X \subset Y$ 
  \attrib{\citep[(104)]{ChierchiaFox2008:The-Grammatical}}
\end{exe}
or
\begin{exe}
\ex \label{bsp:SMH-Partial}  $X$ is preferred to $Y$ iff $X \subset Y$
  and $X$ and $Y$ are readings obtained from applications of
  $\exh(\cdot)$ at exactly the same scope sites, except for
  one. \attrib{\citep[(105)]{ChierchiaFox2008:The-Grammatical}} 
\end{exe}
Since the former is entailed by the latter, we will speak of
\mymark{weak grammaticalism} and \mymark{strong grammaticalism},
depending on whether (\ref{bsp:SMH-Total}) or (\ref{bsp:SMH-Partial})
is used.

@@

Though there seems to be general consensus about such potential
prosodic effects, empirical evidence for this assumption is still
scarce. One notable exception comes from a pilot study reported in
\citet{Frazier08}, in which participants were instructed to read
embedded and non-embedded versions of English \as-sentences. In the
embedded sentences (e.g. {\it All of the students wrote some of the
  official memos}), {\it some} was either capitalized or
non-capitalized. Prior to the experiment, participants were informed
that capitalizing corresponded to contrastive stress. The task
consisted in a forced-choice questionnaire with two given alternatives
(such as {\it All of the students wrote some but not all of the
  official memos} vs. {\it All of the students wrote at least some of
  the official memos}). Interestingly, strengthened interpretations
occurred rather often across all sentences in this task (59 \%), but
were not affected by capitalizing.\fn{Note that, unfortunately, the
  authors do not provide percentages for the embedded conditions
  alone, but only the means across sentence types.}
\citeauthor{Frazier08} concludes that accentuation did not affect
their silent reading data, but raises the possibility that prosodic
effects may have been present at least in a subset of the items (see
footnote~1 in \citet{Frazier08}, p.~330).

>>>>

\usepackage[natbib=true,
            style=authoryear-comp,
            backend=bibtex,
            sortcites=false,
            maxnames=2,
            doi=false,
            url=false]{biblatex}
\bibliography{MyRefGlobal}
% \bibliography{paper}

@@

Consequently, the \mymark{Gricean recipe}
\citep[see][]{Geurts2010:Quantity-Implic} for deriving a simple scalar
inference like that in (\ref{bsp:Plain-SI-Implicature}) from an
utterance of (\ref{bsp:Plain-SI-Target}) is as follows: if the issue
whether Hans solved only some or all of the problems is relevant, then
a cooperative and knowledgable speaker would utter
(\ref{bsp:Plain-SI-Alternative}) if in a position to do so; hence, one
of the most natural explanations of why such a speaker has not uttered
(\ref{bsp:Plain-SI-Alternative}), but only (\ref{bsp:Plain-SI-Target})
is that she is uncertain of whether (\ref{bsp:Plain-SI-Alternative})
is true; but on the assumption that she is knowledgeable (competent,
opinionated, informed \dots) it follows that
(\ref{bsp:Plain-SI-Implicature}) should in fact be true.\fn{We are
  glossing here somewhat swiftly over the more nuanced details of the
  derivation of implicatures targeting the speaker's epistemic state
  \citep[e.g.][]{Gazdar1979:Pragmatics:-Imp,Soames1982:How-Presupposit}. The
  concluding Section~\ref{sec:conclusions} will come back to this in
  more detail.}

@@

Traditionalism's predictions about preferences between literal and
global readings depend on whether the auxiliary assumptions necessary
to derive global readings by the Gricean recipe are plausibly
met. These auxiliary assumptions include relevance of the extra
information provided by the global reading, mutual awareness that the
stronger alternative has been a speaker option, the speaker's
competence about the issue, etc. Normally, traditionalist accounts
would assume that these extra assumptions are met. In that case,
traditionalism would predict that global readings should be preferred
over literal readings. On the other hand, if auxiliary assumptions are
implausible, traditionalism would predict that the literal readings
would be preferred over the global one. Consequently, there are at
least two varieties of traditionalism that, depending on which
additional assumptions we would make, yield slightly different
predictions: \mymark{the strong variety of traditionalism} maintains
that the auxiliary assumptions of the Gricean recipe hold
usually/strongly/unless-completely-intenable and therefore predicts
that global readings are preferred over literal ones; \mymark{the weak
  variety of traditionalism} holds that the auxiliary assumptions are
more fragile and predicts that literal readings are preferred over the
global ones.\fn{We mention for clarity that although weak
  traditionalism does not predict the global reading to be preferred,
  it might still predict an \emph{epistemically weak implicature},
  similar to the global reading, that the speaker is uncertain whether
  the stronger alternative is true. (See also the discussion in
  Section~\ref{sec:conclusions}.)}

Does traditionalism also predict local readings to be available? The
answer is slightly different for \as- and the \es-sentences. For
\es-sentences, traditionalism does not predict that a local reading is
available, at least not as a quantity implicature by conjoining the
literal meaning with a suitable set of negated alternatives
\citep[see][]{GeurtsPouscoulous2009:Embedded-Implic,ChemlaSpector2010:Experimental-Ev}. The
literal and the local reading of \es-sentences are logically
independent. But if $X$ and $Y$ are logically independent
propositions, then there is no proposition $Z$ such that $X$ would be
equivalent to $Y \wedge \neg Z$.

@@

Recently, a lot of evidence based on intuitive judgements and more
theoretical considerations has been discussed in favor of
\mymark{grammaticalism}
\citep[see][]{Chierchia2006:Broaden-Your-Vi,Fox2007:Free-Choice-and,Magri2011:Another-Argumen,Sauerland2012:The-Computation,ChierchiaFox2008:The-Grammatical}. % When
% it comes to the predictions of available readings of \as- and
% \es-sentences, grammaticalism appears like the conjunction of
% traditionalism and conventionalism. But conceptually speaking,
% grammaticalism goes down a road of its own.
Grammaticalism approaches quantity implicatures by postulating a
silent operator that can be variably applied during compositional
computation of a sentence's truth-value
\citep{Chierchia2006:Broaden-Your-Vi}, if necessary multiple times
\citep{Fox2007:Free-Choice-and}. This silent operator is referred to
as $\mathrm{O}(\cdot)$ or $\exh(\cdot)$, because it is assumed to be
similar in effect to the meaning of particle \emph{only} or of the
mechanism of \emph{exhaustive interpretation}
\citep{GroenendijkStokhofThesis1984,Stechowvon-StechowZimmermann1984:Term-Answers-an,Rooijvan-RooijSchulz2013:Exhaustive-Inte,vanRooijSchulz:ExhaustiveInterpretation,Fox2007:Free-Choice-and}. For
our purposes, it is enough to note that $\exh(\cdot)$ is a poly-typed
function that enriches an expression $X$, which need not be a full
proposition, based on a set $\alt(X)$ of (suitable, relevant)
alternatives to $X$ that yields an enriched meaning of the
form:\fn{More sophisticated formulations of exhaustive interpretation
  have been proposed
  \citep[e.g.][]{Schulz2005:A-Pragmatic-Sol,vanRooijSchulz:ExhaustiveInterpretation,Spector2006:Scalar-Implicat,Fox2007:Free-Choice-and}
  but this simple formulation is sufficient for the purposes of this
  paper.}
\begin{exe}
  \ex \label{bsp:Exh-Def} $\exh(X,\alt(X)) = X \bigwedge_{A \in
      \alt(X)} \neg A$.
\end{exe}

@@

The grammatical approach, as described so far, is not yet a
respectable theory. It makes little empirical commitments of
substance. It must be supplemented by a selection principle that
determines which of the many readings made available by the grammar
actually obtain or are preferred. Proponents of grammaticalism are
aware of this and have suggested some tentative disambiguation
strategies. \citet{Magri2011:Another-Argumen}, for example, opts for
obligatory presence of $\exh$ with variability in the set of relevant
alternatives. \todo{verify that this is the proposal} Another proposal
is that readings are preferred in relation to how well they answer the
contextually given question under discussion
\citep[e.g.][]{Fox2007:Free-Choice-and,GualminiHulsey2008:The-Question-An}. Yet
another proposal is to select readings by comparison of logical
strengths, as suggested by the \mymark{strongest meaning hypothesis}
of \citet{DalrympleKanazawa1998:Reciprocal-Expr}
\citep[see][]{FoxSpector:Economy-and-Emb,ChierchiaFox2008:The-Grammatical,Chierchia2012:FC-Nominals-and,Spector2014:Global-Positive}. We
will focus entirely on this latter selection principle, because it
seems to be the most prominent in the literature and because it makes
the clearest predictions. That means that all we will say about
grammaticalism in this paper strictly applies only to grammaticalism +
strength based selection of readings.

\citet{ChierchiaFox2008:The-Grammatical} discuss two concrete variants
of strength-based disambiguation: fix a sentence with propositional
content $S$ with candidate readings $C(S)$, where $C(S)$ contains $S$
and all readings derivable from inserting $\exh(\cdot)$ at suitable
scope sites;\fn{Strictly speaking, we should compare \emph{parses} of
  sentences with $\exh(\cdot)$ at various scope sites with respect to
  the readings that these parses give rise to
  \citep[see][]{ChierchiaFox2008:The-Grammatical}, but we may ignore
  this detail here for ease of exposition.} then define for $X,Y \in
C(S)$ either
\begin{exe}
\ex \label{bsp:SMH-Total} 
  $X$ is preferred to $Y$ iff $X \subset Y$ 
  \attrib{\citep[(104)]{ChierchiaFox2008:The-Grammatical}}
\end{exe}
or
\begin{exe}
\ex \label{bsp:SMH-Partial}  $X$ is preferred to $Y$ iff $X \subset Y$
  and $X$ and $Y$ are readings obtained from applications of
  $\exh(\cdot)$ at exactly the same scope sites, except for
  one. \attrib{\citep[(105)]{ChierchiaFox2008:The-Grammatical}} 
\end{exe}
Since the former is entailed by the latter, we will speak of
\mymark{weak grammaticalism} and \mymark{strong grammaticalism},
depending on whether (\ref{bsp:SMH-Total}) or (\ref{bsp:SMH-Partial})
is used.

@@

Though there seems to be general consensus about such potential
prosodic effects, empirical evidence for this assumption is still
scarce. One notable exception comes from a pilot study reported in
\citet{Frazier08}, in which participants were instructed to read
embedded and non-embedded versions of English \as-sentences. In the
embedded sentences (e.g. {\it All of the students wrote some of the
  official memos}), {\it some} was either capitalized or
non-capitalized. Prior to the experiment, participants were informed
that capitalizing corresponded to contrastive stress. The task
consisted in a forced-choice questionnaire with two given alternatives
(such as {\it All of the students wrote some but not all of the
  official memos} vs. {\it All of the students wrote at least some of
  the official memos}). Interestingly, strengthened interpretations
occurred rather often across all sentences in this task (59 \%), but
were not affected by capitalizing.\fn{Note that, unfortunately, the
  paper does not provide percentages for the embedded conditions
  alone, but only the means across sentence types.}
\citeauthor{Frazier08} concludes that accentuation did not affect the
silent reading data, but raises the possibility that prosodic effects
may have been present at least in a subset of the items (see
footnote~1 in \citet{Frazier08}, p.~330).
